---
title: "Preliminary data exploration"
author: "MostMisidentifiedSpecies team"
date: '2020-05-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Import data for observed bees in greater montreal as of 2020-05-27, 3:30pm

```{r import data using iNatTools package}

# Libraries
library("tidyverse")

# Import data subfiles and reassemble the dataset
for(i in 1:3){
  subFile <- readRDS(paste("../data/",i,"-Can_bees_API_2020-05-27.rds",sep = ""))
  
  if(i == 1) data_api <- subFile
  if(i > 1) data_api <- bind_rows(data,subFile)
}

```

The dataset *data* is comprized of 25899 records and 154 columns/variables.


## Lets remove these casual quality grade observations from data_api to avoid mistakes

```{r clean data}

# Remove casual quality grade observations
data <- data_api[! data_api$quality_grade == "casual",]

```


## Available information on observations

```{r Observation information}

# Column names
colnames(data)

```

154 columns/variables.


### What columns may hold information on ID quality

[1] quality_grade: research, needs_id, casual

identifications: [8] identifications_most_agree: T/F, [10] identifications_most_disagree: T/F, [52] identifications_some_agree: T/F

num_identification: [37] num_identification_agreements, [46] num_identification_disagreements

[13] comments_count

[18] quality_metrics: ???

[44] identifications_count:

[55] identifications

[85] taxon.rank_level

[106] taxon.flag_counts.unresolved

[107] taxon.flag_counts.resolved

[151] user.roles: none, curator, admin

**Possibility to get maverick identifications (Taxon is not a descendant or ancestor of the community taxon. The community does not agree with this identification) from API**


### What information did other studies used?


## Observation identification agreement and quality grade


### Identification agreement

Identification agreement can take 3 values: *most_agree*, *most_disagree*, *some_agree*.


### Quality grade

See https://www.inaturalist.org/pages/help#quality

An obervation is associated with *needs_id* quality grade until it atteins *research* grade or is revised as *casual* quality. It is associated with observations without community (2/3) agreement on a specific ID, that are voted needs more ID, or that is not voted as wild/naturalized if there are at least 10 other observations of a genus or lower in the smallest county-, state-, or country-equivalent place that contains this observation and 80% or more of those observations have been marked as not wild/naturalized.

Observations become *research* quality grade when when more than 2/3 of identifiers agree a on species-level ID or lower. Observations may loose *research* quality grade if the community ID shifts above the species-level. Observations can be *research* grade at the genus level if the community agrees on a genus-level ID and votes that the observation does not need more IDs.

Observations that are not verifiable because they do not have a date, are not georeferenced (i.e. have lat/lon coordinates), have no photos or sounds, and aren't of a captive or cultivated organism, get the *casual* quality grade.


## Test proportion of « research grade observations » / « all observations » and for relationship between identification agreement and quality grade in observations

```{r quality grade vs. identification agreement}

# How many possible identification grades?
unique(data$quality_grade)

# How many observations are associated to respective identifications categories?
length(data[data$identifications_most_agree == T,]$identifications_most_agree)
length(data[data$identifications_most_disagree == T,]$identifications_most_disagree)
length(data[data$identifications_some_agree == T,]$identifications_some_agree)

# Calculate propotion of reasearch grade observations

# # For all observations
all <- length(data[data$quality_grade=="research",]$uuid)/length(data$uuid)

# # For most_agree identifications
agreement <- data[data$identifications_most_agree,]
prop_most_agree <- length(agreement[agreement$quality_grade=="research",]$uuid)/length(agreement$uuid)

# # For most_disagree identifications
disagreement <- data[data$identifications_most_disagree,]
prop_most_disagree <- length(disagreement[disagreement$quality_grade=="research"]$uuid)/length(disagreement$uuid)

# # For some_agree identifications
some_agreement <- data[data$identifications_some_agree,]
prop_some_agree <- length(some_agreement[some_agreement$quality_grade=="research",]$uuid)/length(some_agreement$uuid)

```

Observations have only one of two quality grades, *needs_id* or *reseach*.

Turns out observations identified as *most_agree* and *some_agree* are the same. What is the difference between both identiciations?

There are 367 observations identified as *most_agree* and as *some_agree* in the dataset while only 1 *most_disagree* identification.


### Plot results

```{r pressure}

y <- c(all, prop_most_agree,prop_most_disagree,prop_some_agree)
barplot(y, names.arg = c("all \nn=25490", "most_agree \nn=21312","most_disagree \nn=33","some_agree \nn=21315"), ylab = "Proportion research grade", ylim = c(0,1))

```


~ 65% (16435) of observations received *research* quality grade. Others were identified as *needs_id*.

 ~75% (16435) of observations identified as *most_agree* and *some_agree* and none of the *most_disagree* observations recieved *research* quality grade. The remaining (~25% observations identified as *most_agree* and *some_agree* and 100% of *most_disagree* observations) recieved a *needs_id* quality grade.

It appears some observations are found in neither of the identification categories?!?


## Explore which observations are found in neither of the identification categories

```{r why observation identification not categorized?}

# Which observations are in neither of the identification categories

# # Verify which observations are neither in *most_agree*, *some_agree*, or *most_disagree* categories
missing_obs <- unique(data[! data$uuid %in% agreement$uuid &
                             ! data$uuid %in% some_agreement$uuid &
                             ! data$uuid %in% disagreement$uuid
                           ,]$uuid)

# # Verify all missing observation are captuded
length(missing_obs)

# # View
View(data[data$uuid == missing_obs,])

```


## Select the « not research grade » observation that have more than one identifier to select the observations that are disagreed upon

```{r not research quality grade}

# Select not research quality grade observations that have multiple identifications
notResearchObs <- data[! data$quality_grade == "research" &
                         data$identifications_count > 1,]

# View distribution of identification counts
hist(notResearchObs$identifications_count, xlab = "Number identification count \nn=1943")

```


## Are the non-research grade observations really disagreed upon?

```{r disagreement in non-research quality grade}

# Subseting data.frame for plotting purposes 
library(tidyverse)
notResearchObs_plot <- notResearchObs %>%
                        select(id,
                               identifications_count,
                               num_identification_agreements)
notResearchObs_plot$id <- as.factor(notResearchObs_plot$id)

#Creating a new column corresponding to the percentage of identifiers' agreement 
notResearchObs_plot$PercAgreement <- notResearchObs_plot$num_identification_agreements/notResearchObs_plot$identifications_count

# Plot 
library(ggplot2)
ggplot(notResearchObs_plot, aes(x=id,y=PercAgreement)) + 
    geom_bar(stat="identity") + 
    scale_y_continuous(labels=scales::percent)+ 
    ggtitle("Even for non-research grade observations, \nidentifiers tend to agree on the identifications")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
    ylab("Identifiers agreement")+
    xlab("Observation ID")
    
```


## Research grade vs needs ID 

```{r Quality grade}

ggplot(data=data, aes(x= species_guess, fill= quality_grade))+
  ggtitle("Quality grade for each observation of a specific taxa")+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5))+
  ylab("Count")+
  xlab("Scientific name")
  
```

The plot is not very clear due to the ammount of observations


## **Objective** : select the taxa which have more "needs_id" observation than "research grade" observations

```{r Quality grade, needs_id > research grade}

Bees_CA_Qgrade <- plyr::count(data, c("species_guess","quality_grade"))
# Counts the occurences of quality grade per taxa (scientific name)

library(tidyr)
Bees_CA_Qgrade_wide <- spread(Bees_CA_Qgrade, quality_grade, freq)# Converts to wide format 
Bees_CA_Qgrade_wide[is.na(Bees_CA_Qgrade_wide)] <- 0 # Replaces NA by 0

Bees_Needs_ID <- subset(Bees_CA_Qgrade_wide, Bees_CA_Qgrade_wide$needs_id>Bees_CA_Qgrade_wide$research)
#  Selects he taxa which have more "needs_id" observation than "research grade" observations

library(reshape2)
Bees_Needs_ID_long <- melt(Bees_Needs_ID)# Convert to wide format for plotting

# Plot
ggplot(data=Bees_Needs_ID_long, aes(x = species_guess,y = value, fill = variable))+
  ggtitle("Taxa for which there are more needs_id observations than research grade")+
  geom_bar(stat="identity",position="stack")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
        legend.position = "bottom")+
  scale_fill_discrete("Quality grade")+
  ylab("Count")+
  xlab("Scientific name")

``` 

Ohhh interesting results!!!

Although this still lacks clarity, it seems that many taxa in this plot are not identified to species level.
