---
title: "glm_modelling_jd"
author: "MostMisidentifiedSpecies Team"
date: '2020-06-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import data}
library(tidyverse)

data<-read.csv("../data/iNatDataset_Normalized_SpeciesCount_SubsetforAnalysis.csv")

```

```{r clean data}
##removing row count column
data<-data[c(-1)]

##removing casual grade observations
data <- data[! data$quality_grade == "casual",]

##removing NAs/RatioMisID>0
finaldata <- data %>% subset(RatioDisagree>0)

##removing observations not ID to species level
finaldata <- finaldata[finaldata$taxon.rank == "species",]
  
```

Data is all cleaned up and ready to be explored.

```{r correlation matrix on parameters of interest}
##corrmatrix
library(Hmisc)
library(corrplot)

s_corrmatrix <- cor(finaldata[, colnames(finaldata) %in% c("species_count","user.identifications_count", "taxon.observations_count", "taxon.taxon_changes_count", "comments_count", "RatioDisagree")], method = "spearman")

round(s_corrmatrix, 3)
corrplot(s_corrmatrix)

s_corr <- rcorr(as.matrix(s_corrmatrix, type = c("spearman")))
s_corr$r
s_corr$P

p_corrmatrix <- cor(finaldata[, colnames(finaldata) %in% c("species_count","user.identifications_count", "taxon.observations_count", "taxon.taxon_changes_count", "comments_count")], method = "pearson")

round(p_corrmatrix, 3)
corrplot(p_corrmatrix)

p_corr <- rcorr(as.matrix(p_corrmatrix, type = c("pearson")))
p_corr$r
p_corr$P

##chi square test for categorical parameters
chisq.test(finaldata$location, finaldata$observed_on)
chisq.test(finaldata$location, finaldata$quality_grade)
chisq.test(finaldata$location, finaldata$taxon.name)
chisq.test(finaldata$observed_on, finaldata$quality_grade)
chisq.test(finaldata$observed_on, finaldata$taxon.name)
chisq.test(finaldata$quality_grade, finaldata$taxon.name)

```

We can see from the Spearman's that species_count + taxon.taxon_changes_count are highly correlated. species_count is also correlated with taxon.observations_count (p=0.089).

We can see from the Chi-squared tests that location and observed_on are highly correlated and quality_grade and taxon.name are also highly correlated.

Based on these correlation tests, we will use species_count, location, and taxon.name.

```{r building glm models}
##creating dataframe with only parameters used in the model
library(dplyr)
mod_data <- dplyr::select(finaldata, RatioDisagree, species_count, location, taxon.name, user.identifications_count)

mod_data <- mod_data %>% mutate(z_score = scale(value))# Z score the variables
                          
mod_data <- mod_data %>% mutate_at(c(1, 2, 3, 5), funs(c(scale(mod_data, center = TRUE, scale = TRUE)))) # Z score the variables
mod_data<- na.omit(mod_data) # remove NAs

model <- glm(RatioDisagree ~ comments_count + user.identifications_count + identifications_count + Freq + taxon.taxon_schemes_count, 
              data = mod_data)
summary(model)  
  
##building glm w/ hypothesized, uncorrelated parameters, pairwise by taxon.name?
histogram(finaldata$RatioDisagree)
qqnorm(finaldata$RatioDisagree)

glm.test <- glm(RatioDisagree ~ species_count + location + taxon.name, data = finaldata)
summary(glm.test)

###aggregate by species
species <- finaldata %>%
  #filter(taxon.rank == "species") %>%
  group_by(taxon.name) %>%
  #add_tally() %>%
  dplyr::summarize(disagreementScore = mean(num_identification_disagreements)/mean(identifications_count), 
            num_identification_disagreements=sum(num_identification_disagreements), 
            num_identification_agreements=sum(num_identification_agreements), 
            identifications_count=sum(identifications_count),
            comments_count=sum(comments_count),
            user.identifications_count=sum(user.identifications_count),
            genus=genus[1],
            species_count=species_count[1],
            n()) %>%
  filter(!is.na(.$disagreementScore))

species$NoObs <- species$`n()`

model <- glm(disagreementScore ~ comments_count + user.identifications_count + identifications_count + species_count + NoObs, data = species)

##looking at results
summary(model)  
Anova(model)

summ(model)

```

```{r creating glm result figure}

##creating forest plot - or similar
library(jtools)
library(ggstance)
library(broom.mixed)

plot_summs(model, scale = TRUE)

```